Formal Analysis of Rate-Monotonic Scheduling Implementation via Real-Time Maude

### 2015-10-07 RMS/submission.tex
finish writing the experiment results

% This is LLNCS.DOC the documentation file of
% the LaTeX2e class from Springer-Verlag
% for Lecture Notes in Computer Science, version 2.4
\documentclass{llncs}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}
%\usepackage{llncsdoc}

@@

In this section, we analyze our model of the RMS implementation under
several different scenarios, containing both realistic and
experimental ones.

The main objective is to verify the schedubility of a given set of
periodic tasks. Therefore we define an atomic proposition
\verb|taskTimeout| to hold if there exists an \verb|error| appearing
in the $taskList$ of the current state:
\begin{verbatim}
  op taskTimeout : -> Prop [ctor] .
  eq {L T STS HW ISRC} |= taskTimeout = containError(L) .
\end{verbatim}

\subsection{Properties Formalism}

>>>>

% This is LLNCS.DOC the documentation file of
% the LaTeX2e class from Springer-Verlag
% for Lecture Notes in Computer Science, version 2.4
\documentclass{llncs}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{enumerate}
\usepackage{epic}
\usepackage{pstricks}
%\usepackage{llncsdoc}

@@

In this section, we analyze our model of the RMS implementation under
several different scenarios, including both realistic and experimental
ones. Notice that from any (reasonable) given initial state, the
number of reachable states is finite, thanks to the upperbound given
to the $timer$, which provides the potential for applying the untimed
model-checker.

\subsection{Properties Verified}
The main objective is to verify the schedulability of a given set of
periodic tasks. Therefore we define an atomic proposition
\verb|taskTimeout| to hold if there exists an \verb|error| appearing
in the $taskList$ of the current state:
\begin{verbatim}
  op taskTimeout : -> Prop [ctor] .
  eq {L T STS HW ISRC} |= taskTimeout = containError(L) .
\end{verbatim}
where \verb|containError| returns \verb|true| if there is an object
\verb|error| existing in $L$. Then our desired
property--schedulability--can be formalized as the temporal logic
formula \verb|[](~taskTimeout)|. As the property is not
\emph{clocked}, given an initial state \verb|init|, the following
untimed model-checking command returns \verb|true| if our desired
schedulability property holds \emph{with no time limit}; otherwise 
a trace showing a counterexample is provided.
\begin{verbatim}
  (mc init |=u [](~taskTimeout) .)
\end{verbatim}

The other objective is to verify the correctness of the
implementation, in the sense that the periodic tasks are scheduled
exactly with respect to their priorities. The atomic proposition
\verb|scheduleCorrectly| is defined to hold if the running periodic
task is the one requested to be run and with the highest priority:
\begin{verbatim}
  op scheduleCorrectly : -> Prop [ctor] .
  ceq {L T STS HW ISRC} |= scheduleCorrectly 
        = (not containError(L))
          and (if ID :: MaybeNat then shouldRun(ID, L)
               else true fi)
      if ID := (HW).getPc .
\end{verbatim}
where \verb|shouldRun(ID, L)| returns \verb|true| if the task
identified by \verb|ID|, probably \verb|none|, is the one possessing
the highest priority among those whose status is \verb|READY|,
\verb|RUNNING| or \verb|INTERRUPT|. Note that \verb|scheduleCorrectly|
requires no \verb|error| popping up, since our defined correctness
property makes no sense in the implementation once some task misses
its deadline. The correctness property is formalized as the temporal
logic formula \verb|[]scheduleCorrectly|, and can be verified by the
following untimed model-checking command provided an initial state
\verb|init|, which returning \verb|true| indicates the correctness
property holds with no time limit.
\begin{verbatim}
  (mc init |=u []scheduleCorrectly .)
\end{verbatim}

\subsection{Scenarios}
We use the following setting for our verification, which is from the 
statistics provided by our industrial partner:
\begin{itemize}
\item The interrupt cycle is $5ms$.
\item The scheduling time ranges from $5{\mu}s$ to $9{\mu}s$, and the
switching time ranges from $2{\mu}s$ to $4{\mu}s$.
\item The initial state is with empty stack, empty \verb|pc|, cleared 
\verb|mask| and cleared \verb|ir|.
\end{itemize}

We have analyzed our model/implementation in ten different scenarios,
including both realistic ones provided by our industrial partner and
experimental ones designed by ourselves, four of them are described
below:
\begin{itemize}
\item Scenario (i) with two tasks: one having period $5ms$ and
  possible finishing interval $[2.4ms, 3ms]$, and the other having
  period $25ms$ and possible finishing interval $[6.1ms, 7ms]$.
\item Scenario (ii) with two tasks: one having period $5ms$ and
  possible finishing interval $[1.8ms, 2ms]$, and the other having
  period $25ms$ and possible finishing interval $[2.1ms, 2.3ms]$.
\item Scenario (iii) with three tasks: the first having period $5ms$
  and possible finishing interval $[2.5ms, 2.7ms]$, the second having
  period $10ms$ and possible finishing interval $[1.5ms, 2ms]$, and
  the third having period $25ms$ and possible finishing interval
  $[2.6ms, 3ms]$.
\item Scenario (iv) with three tasks: the first having period $5ms$
  and possible finishing interval $[2.2ms, 2.5ms]$, the second having
  period $10ms$ and possible finishing interval $[1.4ms, 1.5ms]$, and
  the third having period $15ms$ and possible finishing interval
  $[4.3ms, 4.5ms]$.
\end{itemize}

Choosing the \emph{maximal time sampling strategy}, the results of the
model checking show that both the schedulability and the correctness
properties hold in Scenarios (i-iii), but not in Scenario (iv). For
Scenario (iv), one counterexample of the schedulability is pictured in
Figure~\ref{f:counterexample}, where the third task (task$2$) misses
its deadline at the time $15ms$ that is marked by the tiny red
triangle. And the correctness fails in Scenario (iv) because of the
failure of schedulability.

\begin{figure}[h]
\begin{center}
\begin{picture}(300,110)(0,-10)
\thicklines
\put(0,0){\vector(1,0){320}}
\put(305,5){time(${\mu}s$)}
\thinlines
\put(0,0){\line(0,1){110}}
\put(-25,97){task$0$}
\put(-25,77){task$1$}
\put(-25,57){task$2$}
\put(-44,37){scheduling}
\put(-40,17){switching}
\put(-2,-8){$0$}

\dashline[30]{3}(0,100)(250,100)
\dashline[30]{3}(0,80)(280,80)
\dashline[30]{3}(0,60)(300,60)
\dashline[30]{3}(0,40)(210,40)
\dashline[30]{3}(0,20)(285,20)

\dashline[30]{3}(10,-10)(10,100)
\thicklines
\textcolor{red}{\put(0,40){\line(1,0){10}}}
\put(8,-18){$9$}

\thinlines
\dashline[30]{3}(50,0)(50,100)
\thicklines
\textcolor{blue}{\put(10,100){\line(1,0){40}}}
\put(35,-8){$2509$}

\thinlines
\dashline[30]{3}(55,-10)(55,80)
\thicklines
\textcolor{red}{\put(50,20){\line(1,0){5}}}
\put(46,-18){$2513$}

\thinlines
\dashline[30]{3}(80,0)(80,80)
\thicklines
\textcolor{blue}{\put(55,80){\line(1,0){25}}}
\put(65,-8){$4013$}

\thinlines
\dashline[30]{3}(85,-10)(85,60)
\thicklines
\textcolor{red}{\put(80,20){\line(1,0){5}}}
\put(76,-18){$4017$}

\thinlines
\dashline[30]{3}(100,0)(100,60)
\thicklines
\textcolor{blue}{\put(85,60){\line(1,0){15}}}
\put(90,-8){$5000$}

\thinlines
\dashline[30]{3}(110,-10)(110,100)
\thicklines
\textcolor{red}{\put(100,40){\line(1,0){10}}}
\put(101,-18){$5009$}

\thinlines
\dashline[30]{3}(150,0)(150,100)
\thicklines
\textcolor{blue}{\put(110,100){\line(1,0){40}}}
\put(135,-8){$7509$}

\thinlines
\dashline[30]{3}(155,-10)(155,60)
\thicklines
\textcolor{red}{\put(150,20){\line(1,0){5}}}
\put(146,-18){$7513$}

\thinlines
\dashline[30]{3}(200,0)(200,60)
\thicklines
\textcolor{blue}{\put(155,60){\line(1,0){45}}}
\put(185,-8){$10000$}

\thinlines
\dashline[30]{3}(210,-10)(210,100)
\thicklines
\textcolor{red}{\put(200,40){\line(1,0){10}}}
\put(198,-18){$10009$}

\thinlines
\dashline[30]{3}(250,0)(250,100)
\thicklines
\textcolor{blue}{\put(210,100){\line(1,0){40}}}
\put(230,-8){$12509$}

\thinlines
\dashline[30]{3}(255,-10)(255,80)
\thicklines
\textcolor{red}{\put(250,20){\line(1,0){5}}}
\put(243,-18){$12513$}

\thinlines
\dashline[30]{3}(280,0)(280,80)
\thicklines
\textcolor{blue}{\put(255,80){\line(1,0){25}}}
\put(260,-8){$14013$}

\thinlines
\dashline[30]{3}(285,-10)(285,60)
\thicklines
\textcolor{red}{\put(280,20){\line(1,0){5}}}
\put(272,-18){$14017$}

\thinlines
\dashline[30]{3}(300,0)(300,60)
\thicklines
\textcolor{blue}{\put(285,60){\line(1,0){15}}}
\put(289,-8){$15000$}

\textcolor{red}{\put(300,60){\vector(-1,0){0}}}

\end{picture}
\end{center}
\caption{An Counterexample of Scenario (iv).}
\label{f:counterexample}
\end{figure}
